The future age structure of scientific disciplines

### 2025-04-16 age_structure_forecasts.qmd
Formatting

Retirement data is sourced from the *Retirement and Retirement Intentions* dataset (Catalogue 6238) for the 2022–2023 financial year [@ABS6238]. It provides insights into expected retirement ages for workers in age groups ranging from 45 to 70 across various industries.

@@

Younger and older age groups within some disciplines were particularly affected by this volatility, especially when sample sizes were small. For example, in a group of five 15–19-year-olds, a single individual's employment status change could significantly shift the participation rate. Similarly, fluctuations in smaller cohorts of older scientists impacted participation rates. Averaging the rates from 2016 and 2021 helps mitigate this volatility and provides a more stable estimate for 2006 and 2011. Any discrepancies in participation rates are minimal due to the small cohort sizes, making this a more reliable approach than relying on a single year.

@@

```{r}
#| label: fig-cohort-interpolation
#| eval: true
#| fig-cap: Interpolated number of scientists in Australia, 2006–2021, using linear cohort interpolation.
tar_read(fig3)
```

@@

```{r}
#| label: fig-net-migration-disciplines
#| eval: true
#| fig-cap: Estimated net migration by age across scientific disciplines (2006–2020). Each panel represents a different discipline, showing net migration trends across age groups.
#| fig-height: 6
tar_read(fig22)
```

@@

-   **Physics and Astronomy** *(Aging Workforce):* The 35–45 age group shows a higher average than the previous 2025 cohort. There is a growing number of scientists aged 45 to 60, and more individuals aged 60+ are remaining in the workforce longer. These patterns indicate a gradual transition toward an older age structure, reflecting an aging workforce.
-   **Mathematical Sciences** *(Expansive Workforce):* Workforce growth is evident across a broad range of ages, particularly among individuals aged 25 to 70. The most significant increase occurs before 60, though there is also a clear rise in older workers (60+) remaining in the workforce rather than retiring. Overall, workforce growth is strong and evident across age groups.
-   **Chemical Sciences** *(Stationary Workforce):* The average workforce size remains stable, with a relatively narrow prediction interval, indicating a predictable pattern of stability over time. While there is a minor increase in the number of workers aged 70, this change is not substantial enough to suggest a broader shift in workforce dynamics.
-   **Earth Sciences** *(Aging Workforce):* Early-career workforce projections are more uncertain relative to its size. A decline in the working population is observed in the mid-thirties, followed by an increase throughout the forties, though both remain modest. While the overall workforce remains relatively stable, these patterns indicate a slow shift toward an older age distribution.
-   **Biological Sciences** *(Aging Workforce):* The number of workers aged 50 to 65 is increasing, indicating growth in later-career participation.
-   **Other Natural and Physical Sciences** *(Expansive Workforce):* Workforce growth is strong across ages 30 to 70, characteristic of an expanding workforce.
-   **Natural and Physical Sciences (n.f.d.)** *(Aging Workforce):* Growth is primarily concentrated in the 55–70 age range, indicating delayed retirement and an aging workforce.

>>>>

Retirement data is sourced from the *Retirement and Retirement Intentions* dataset (Catalogue 6238) for the 2022--2023 financial year [@ABS6238]. It provides insights into expected retirement ages for workers in age groups ranging from 45 to 70 across various industries.

@@

Younger and older age groups within some disciplines were particularly affected by this volatility, especially when sample sizes were small. For example, in a group of five 15--19-year-olds, a single individual's employment status change could significantly shift the participation rate. Similarly, fluctuations in smaller cohorts of older scientists impacted participation rates. Averaging the rates from 2016 and 2021 helps mitigate this volatility and provides a more stable estimate for 2006 and 2011. Any discrepancies in participation rates are minimal due to the small cohort sizes, making this a more reliable approach than relying on a single year.

@@

```{r}
#| label: fig-cohort-interpolation
#| eval: true
#| fig-cap: Interpolated number of scientists in Australia, 2006--2021, using linear cohort interpolation.
tar_read(fig3)
```

@@

```{r}
#| label: fig-net-migration-disciplines
#| eval: true
#| fig-cap: Estimated net migration by age across scientific disciplines (2006--2020). Each panel represents a different discipline, showing net migration trends across age groups.
#| fig-height: 6
tar_read(fig22)
```

@@

-   **Physics and Astronomy** *(Aging Workforce):* The 35--45 age group shows a higher average than the previous 2025 cohort. There is a growing number of scientists aged 45 to 60, and more individuals aged 60+ are remaining in the workforce longer. These patterns indicate a gradual transition toward an older age structure, reflecting an aging workforce.
-   **Mathematical Sciences** *(Expansive Workforce):* Workforce growth is evident across a broad range of ages, particularly among individuals aged 25 to 70. The most significant increase occurs before 60, though there is also a clear rise in older workers (60+) remaining in the workforce rather than retiring. Overall, workforce growth is strong and evident across age groups.
-   **Chemical Sciences** *(Stationary Workforce):* The average workforce size remains stable, with a relatively narrow prediction interval, indicating a predictable pattern of stability over time. While there is a minor increase in the number of workers aged 70, this change is not substantial enough to suggest a broader shift in workforce dynamics.
-   **Earth Sciences** *(Aging Workforce):* Early-career workforce projections are more uncertain relative to its size. A decline in the working population is observed in the mid-thirties, followed by an increase throughout the forties, though both remain modest. While the overall workforce remains relatively stable, these patterns indicate a slow shift toward an older age distribution.
-   **Biological Sciences** *(Aging Workforce):* The number of workers aged 50 to 65 is increasing, indicating growth in later-career participation.
-   **Other Natural and Physical Sciences** *(Expansive Workforce):* Workforce growth is strong across ages 30 to 70, characteristic of an expanding workforce.
-   **Natural and Physical Sciences (n.f.d.)** *(Aging Workforce):* Growth is primarily concentrated in the 55--70 age range, indicating delayed retirement and an aging workforce.
